
\section{Background}

\subsection{PyPy}

\subsubsection{A quick overview}

The PyPy project is basically two things:

\begin{enumerate}
\item the RPython toolchain, written in Python, is a set of compiler tools for 
programs written in RPython.
\item and an implementation of Python using these tools.
\end{enumerate}

In this paper PyPy refers to former.

The basic concept of PyPy is to use a high-level language to allow for rapid
development of interpreters for a variety of platforms. By implementing a compiler
for RPython, interpreters for other languages can be written in RPython and 
compiled to any platform supported by the PyPy toolchain. Supported platforms include
CLI and JVM. \cite{ancona2007rpython}

PyPy uses the meta-programming argument; if a VM (virtual machine) can be written
at a level of abstraction high enough, then it should be possible to automatically translate 
this VM to other lower-level platforms. This is what PyPy does. \cite{pypy}

\subsubsection{RPython}

RPython (Restricted Python) is a restricted proper subset of Python that it 
is possible to perform type inference on. This means that it can be 
translated to efficient C code, and it enables easy analyzis 
as well as efficient compilation. This also means that RPython code can be
run and debugged by Python interpreters, like CPython. \cite{ancona2007rpython}

\subsubsection{Just-in-time compilation}

The JIT compiler is the reason why PyPy is able to compete with other language implementations
on speed. Or rather, it's meta-tracing JIT. The JIT is implemented
for the RPython compiler, but through a set of compiler hints, it is able to trace the 
execution of the application interpreted by the RPython program.

\subsubsection{Haskell-Python}

Haskell-Python is an interpreter for a subset of the Haskell language, called Core'.
We call it Core' (Core marked) here because it does not directly correspond to 
the Core language
used by GHC (the glasgow haskell compiler). Haskell-Python is written in RPython,
and is compiled into a JIT compiler using the RPython toolchain. Our goal for this
project is to extend Haskell-Python to use GHC as a frontend for compilation 
of Haskell programs.

\subsection{GHC}

GHC (Glasgow Haskell Compiler) is a compiler for the Haskell programming language.
\emph{Haskell} is a \emph{strongly-typed non-strict purely-functional} 
programming language, it will not be described in any detail here, since 
Haskell is not the language we focus on. See \cite{hudak1992report}
for an introduction to Haskell. 

Core is an intermediate language used by the Glasgow Haskell Compiler\cite{ghc},
and it is this language we wish to interpret. Core is a desugared version of Haskell, 
things like pattern matching
and list comprehensions are transformed out to simpler constructs.\cite{jones1994compilation}

\subsection{Extcore}

Extcore is a Haskell package for working with GHC's Core language. Among other things,
it implements a parser for External-core, this is the part used from extcore in this project.

\begin{comment}
\subsubsection{Compilation by transformation}

GHC uses a compilation idiom called \emph{compilation by transformation}. The idea is to repeatedly perform 
correctness-preserving transformations to the program. Ideally, these transformations will result
in a semantically equal program that executes more quickly or in less space. Such transformations seem
to fall into two categories:

\begin{itemize} 
\item{\emph{Glamorous transformations}} are global, sophisticated, intellectually satisfying transformations,
sometimes guided by some interesting kind of analyzis.
\item{\emph{Humble transformations}} are small, simple, local transformations. Individually they look very trivial.
\end{itemize}

In the Glasgow Haskell Compiler, all humble transformations are done by the \emph{simplifier}. 
\cite{jones1994compilation} The simplifier repeatedly performs transformations on the Core language.


\subsection{Extcore}

Extcore is a package for working with the External-core format.

\end{comment}
