
\section{External-core}

The Core language is an intermediate language used by GHC.
All syntactic sugar is removed, type checking is performed, pattern matching is
translated into case-expressions (each of witch performs only a single level of
matching) and overloading is resolved.\cite{jones1992implementing} Following
is a description of external-core as presented by \cite{tolmach2010ghc}.

\subsection{External-core language definition}

We use the following semantics to define the Core grammar:

\begin{longtable}{ l c l }

$[$ pat $]$		& :	& optional			\\
$\{$ pat $\}$		& :	& zero or more repetitions	\\
$\{$ pat $\}^{+}$	& :	& one or more repetitions	\\
$pat_{1}|pat_{2}$	& :	& choice			\\

\end{longtable}

\begin{footnotesize}
\begin{longtable}{ r c l r }


\\[0.01in]

\multicolumn{4}{l}{Module}			 \\
\\[0.01in]
$module$	& $ \rightarrow $ 	& \%module $mident$ $\{$ $tdefg$ ; $\}$ $\{$ $vdefg$ ; $\}$				&			\\
\\[0.01in]

\multicolumn{4}{l}{Type defn.}			 \\
\\[0.01in]
$tdefg$ 	& $ \rightarrow $	& \%data $qtycon$ $\{$ $tbind$ $\}$  = $\{$ $[$ $cdef$ $\{$ ; $cdef$ $\}$ $]$ $\}$	& algebraic type	\\
		& $ | $			& \%newtype $qtycon$ $qtycon$ $\{ tbind \}$ = $ty$					& newtype		\\
\\[0.01in]

\multicolumn{4}{l}{Constr. defn.}			 \\
\\[0.01in]
$cdef$		& $ \rightarrow $	& $qdcon$ $\{$ @ $tbind$ $\}$ $\{$ $aty$ $\}^{+}$ 					& 			\\
\\[0.01in]

\multicolumn{4}{l}{Value defn.}			 \\
\\[0.01in]
$vdefg$		& $ \rightarrow $	& \%rec $\{$ $vdef$ $\{$ ; $vdef$ $\}$ $\}$						& recursive		\\
		& $ | $			& $vdef$										& non-recursive		\\
$vdef$ 		& $ \rightarrow $	& $qvar$ :: $ty$ = $exp$								& 			\\
\\[0.01in]

\multicolumn{4}{l}{Atomic expr.}			 \\
\\[0.01in]
$aexp$		& $ \rightarrow $	& $qvar$										& variable		\\
		& $ | $			& $qdcon$										& data constructor	\\
		& $ | $			& $lit$											& literal		\\
		& $ | $			& ( $exp$ ) 										& nested expr.		\\
\\[0.01in]

\multicolumn{4}{l}{Expression}			 \\
\\[0.01in]
$exp$		& $ \rightarrow $	& $aexp$										& atomic expr.		\\
		& $ | $			& $aexp$ $\{$ $arg$ $\}^{+}$ 								& application		\\
		& $ | $			& $\backslash$ $\{$ $binder$ $\}$ -$>$ $exp$						& abstraction		\\
		& $ | $			& \%let	$vdefg$ \%in $exp$								& local definition	\\
		& $ | $			& \%case ( $aty$ ) $exp$ \%of $vbind$ $\{$ $alt$ $\{$ ; $alt$ $\}$ $\}$			& case expr.		\\
		& $ | $			& \%cast $exp$ $aty$									& type coercion		\\
		& $ | $			& \%note "  $\{$ $char$ $\}$ " $exp$							& expression note	\\
		& $ | $			& \%external ccal " $\{$ $char$ $\}$ " $aty$						& external reference	\\
		& $ | $			& \%dynexternal ccal $aty$								& external reference (dynamic)	\\
		& $ | $			& \%label " $\{$ $char$ $\}$ "								& external label	\\
\\[0.01in]

\multicolumn{4}{l}{Argument}			 \\
\\[0.01in]
$arg$		& $ \rightarrow $	& @ $aty$										& type argument		\\
		& $ | $			& $aexp$										& value argument	\\
\\[0.01in]

\multicolumn{4}{l}{Case alt}			 \\
\\[0.01in]
$alt$		& $ \rightarrow $	& $qdcon$ $\{$ @ $tbind$ $\}$ $\{$ $vbind$ $\}$ -$>$ $exp$				& constructor alternative \\
		& $ | $			& $lit$ -$>$ $exp$									& literal alternative 	\\
		& $ | $			& \%\_ -$>$ $exp$									& default alternative	\\
\\[0.01in]

\multicolumn{4}{l}{Binder}			 \\
\\[0.01in]
$binder$	& $ \rightarrow $	& @ $tbind$										& type binder		\\
		& $ | $			& $vbind$										& value binder		\\
\\[0.01in]

\multicolumn{4}{l}{Type binder}			 \\
\\[0.01in]
$tbind$		& $ \rightarrow $	& $tyvar$										& implicit of kind * 	\\
		& $ | $			& ( $tyvar$ :: $kind$ )									& explicitly kinded	\\
\\[0.01in]

\multicolumn{4}{l}{Value binder}			 \\
\\[0.01in]
$vbind$		& $ \rightarrow $	& ( $var$ :: $ty$ )									& \\
\\[0.01in]

\multicolumn{4}{l}{Literal}			 \\
\\[0.01in]
$lit$		& $ \rightarrow $	& ( $[$-$]$ $\{$ $digit$ $\}^{+}$ :: $ty$ )						& integer 		\\ 
		& $ | $			& ( $[$-$]$ $\{$ $digit$ $\}^{+}$ \% $\{$ $digit$ $\}^{+}$ :: $ty$ )			& rational		\\
		& $ | $			& ( ' $char$ ' :: $ty$ )								& character		\\
		& $ | $			& ( " $\{$ $char$ $\}$ " :: $ty$ )							& string		\\
\\[0.01in]

\multicolumn{4}{l}{Character}			 \\
\\[0.01in]
$char$		& $ \rightarrow $	& \multicolumn{2}{l}{Any ASCII character in range 0x20-0x7E except 0x22, 0x27, 0x5c}			 \\
		& $ | $			& $\backslash$x $hex$ $hex$								& ASCII code escape sequence \\
$hex$		& $ \rightarrow $	& 0 $|$ ... $|$ 9 $|$ a $|$ ... f							& \\
\\[0.01in]

\multicolumn{4}{l}{Atomic type}			 \\
\\[0.01in]
$aty$		& $ \rightarrow $	& $tyvar$										& type variable 	\\
		& $ | $			& $qtycon$										& type constructor	\\
		& $ | $ 		& ( $ty$ )										& nested type 		\\
\\[0.01in]

\multicolumn{4}{l}{Basic type}			 \\
\\[0.01in]
$bty$		& $ \rightarrow $	& $aty$											& atomic type		\\
		& $ | $			& $bty$ $aty$										& type application	\\
		& $ | $			& \%trans $aty$ $aty$									& transitive coercion 	\\
		& $ | $			& \%sym	$aty$										& symmetric coercion	\\
		& $ | $			& \%unsafe $aty$ $aty$									& unsafe coercion	\\
		& $ | $			& \%left $aty$										& left coercion		\\
		& $ | $			& \%right $aty$										& right coercion	\\
		& $ | $			& \%inst $aty$ $aty$									& instantiation coercion \\
\\[0.01in]

\multicolumn{4}{l}{Type}			 \\
\\[0.01in]
$ty$		& $ \rightarrow $	& $bty$											& basic type 		\\
		& $ | $			& \%forall $\{$ $tbind$ $\}^{+}$ . $ty$							& type abstraction	\\
		& $ | $			& $bty$ -$>$ $ty$									& arrow type construction \\
\\[0.01in]

\multicolumn{4}{l}{Atomic kind}			 \\
\\[0.01in]
$akind$		& $ \rightarrow $	& $*$											& lifted kind 		\\
		& $ | $			& \#											& unlifted kind 	\\
		& $ | $			& ?											& open kind 		\\
		& $ | $			& $bty$ :=: $bty$									& equality kind 	\\
		& $ | $			& ( $kind$ ) 										& nested kind 		\\
\\[0.01in]

\multicolumn{4}{l}{Kind}			 \\
\\[0.01in]
$kind$		& $ \rightarrow $	& $akind$										& atomic kind		\\
		& $ | $			& $akind$ -$>$ $kind$									& arrow kind		\\
\\[0.01in]

\multicolumn{4}{l}{Identifier}			 \\
\\[0.01in]
$mident$	& $ \rightarrow $	& $pname$ : $uname$									& module		\\
$tycon$		& $ \rightarrow $	& $uname$										& type constr.		\\
$qtycon$	& $ \rightarrow $	& $mident$ . $tycon$									& qualified type constr.\\
$tyvar$		& $ \rightarrow $	& $lname$										& type variable		\\
$dcon$		& $ \rightarrow $	& $uname$										& data constr.		\\
$qdcon$		& $ \rightarrow $	& $mident$ . $dcon$									& qualified data constr.\\
$var$		& $ \rightarrow $	& $lname$										& variable		\\
$qvar$		& $ \rightarrow $	& $[$ $mident$ . $]$ $var$								& optionally qualified variable\\
\\[0.01in]

\multicolumn{4}{l}{Name}			 \\
\\[0.01in]
$lname$		& $ \rightarrow $	& $lower$ $\{$ $namechar$ $\}$								& \\
$uname$		& $ \rightarrow $	& $upper$ $\{$ $namechar$ $\}$								& \\
$pname$		& $ \rightarrow $	& $\{$ $namechar$ $\}^{+}$								& \\
$namechar$	& $ \rightarrow $	& $lower$ $|$ $upper$ $|$ $digit$							& \\
$lower$		& $ \rightarrow $	& a$|$b$|$...$|$z$|$\_									& \\
$upper$		& $ \rightarrow $	& A$|$B$|$...$|$Z$|$									& \\
$digit$		& $ \rightarrow $	& 0$|$1$|$...$|$9									& \\
\\[0.01in]

\end{longtable}
\end{footnotesize}



\subsection{Informal semantics of Core}

Core resembles a explicitly-typed polymorphic lambda-calculus ($F_{w}$), with some additions,
local let bindings, algebraic type definitions, constructors, case-expressions, primitive types,
literals and operators.\cite{tolmach2010ghc}

\subsubsection{Program organization and modules}

Programs represented in Core are organized into modules corresponding directly to source-level
Haskell modules. A module identifier (\emph{mident}) consists of a \emph{package name} followed
by a module name. 

Each module may contain each of the following top-level declarations:
\begin{itemize}
\item{Algebraic datatype declarations:} each defining a type constructor and one or more data
constructors.
\item{Newtype declarations:} corresponding to Haskell newtype declarations, each defining a 
type constructor and a coercion name.
\item{Value declarations:} defining the types and values of top-level variables.
\end{itemize}


\cite{tolmach2010ghc}


\subsubsection{Namespaces}

There are five distinct namespaces:
\begin{enumerate}

\item module identifiers (\emph{mident})
\item type constructors (\emph{tycon})
\item type variables (\emph{tyvar})
\item data constructors (\emph{dcon})
\item term variables (\emph{var})
\end{enumerate}

A variable (type or term) may have multiple definitions within a module. However, they
never shadow one another, the scope of the definition of a variable never contain a
redefinition of the same variable. Type variables may be "shadowed". Thus if a variable has
multiple definitions, they must be local (let-bound).\cite{tolmach2010ghc}

\subsubsection{Types and Kinds}

\paragraph{Types}

\paragraph{Coercions}

\paragraph{Kinds}

\paragraph{Lifted and unlifted types}

\paragraph{Type constructors; base kinds and higher kinds}

\paragraph{Type synonyms and type equivalence}

\subsubsection{Algebraic data types}


\subsubsection{Newtypes}


\subsubsection{Expression forms}


\subsubsection{Expression evaluation}


\subsection{Primitive module}

