

\chapter{Challenges}



\section{Building Haskell libraries}

One of the main challenges of the project was to create External-Core for
the Haskell libraries. This was supposed to be simple, however, due to some GHC
bugs, it was not.

\subsection{Using GHC's '-ext-core' flag}

Not successfull due to bugs in GHC. Milestone set to 7.6.1 (currently at 7.4.1)

\begin{itemize}
\item Panic on generating Core code. (http://hackage.haskell.org/trac/ghc/ticket/5844) 
\item Compiling base with cabal fails. (http://hackage.haskell.org/trac/ghc/ticket/3103)
\end{itemize}

\subsection{Using Cabal to parse '.cabal' file}

Need to translate Cabal information into something understandable by
the GHC API. How do we do that?

\subsection{Interface as GHC}

Create a "fake ghc" program using the GHC API and write simple functionality to 
dump a External-core-like format.

Unfortunately, this is unlikely to work any better, due to wired-in stuff in the
Base package. (Unable to build base independently of GHC)

\subsection{Ignore the problem, implement libraries in Python}

Currently the most attractive solution. But I don't think it's very nice.
The dependencies in external-core programs are complex. Implementing a
"helloworld" program pulls in all kinds of dependencies. Our problem is to 
that we can't implement the functionality at too high a level, and we can't
implement it at too low a level.

Unfortunately, this is also unlikely to work any better, due to wired-in stuff in the
Base package. (Unable to build base independently of GHC)

We can in no way expect to implement all the functionality on our own.

\subsubsection{Perform a dependency check?}

\begin{itemize}
\item Build an AST. 
\item Prune all functionality our program does not depend on.
\item Dump dependency graph.
\item Implement top-most unimplemented functionality. (PROBLEM, we don't know what dependencies a module have without seeing the External-Core of the module).
\end{itemize}
